Climate indicators

### 2023-02-28 DRAFT_climate_indicators.Rmd
parallel cores

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stars)
library(sf)
library(tmap)
library(lubridate)
library(ggpubr)
library(microbenchmark)
library(tictoc)
library(ncmeta)
dir <- substr(getwd(), 1,2)

@@

myFiles <- list.files(path2, pattern=".nc$",full.names = T)[1:2]
real_temp_summer <- NULL

@@

  tic("init")
  temp <- stars::read_ncdf(paste(myFiles[i]), var="tg", proxy=F,
                           ncsub = cbind(start = c(1, 1, julian_start), 
                              count = c(NA, NA, step)))
  year_temp <- year(st_get_dimension_values(temp, "time")[1])
  #start <- ym(paste(year_temp, start_month, sep="-"))
  #end <-  ym(paste(year_temp, end_month, sep="-"))
  #myInterval <- interval(start, end)
  #myInterval <- interval(start, start+1)
  lookup <- setNames("mean", paste0("v_", year_temp))
  st_crs(temp) <- 25833
  toc()

  tic("filter and st_apply")
  temp <- temp %>%
    #filter(time %within% myInterval) %>%
    st_apply(1:2, mean) %>%
    rename(all_of(lookup)) 
  toc()

@@

@@

```{r}
tif = system.file("tif/L7_ETMs.tif", package = "stars")
t1 = as.Date("1970-05-31")
restProxy = read_stars(tif,
               proxy = T)
class(restProxy)
restProxy2 <- st_as_stars(restProxy)
class(restProxy2)
dim(restProxy)

restProxy3 <- restProxy %>% filter(band==1)
dim(restProxy3)
restProxy4 <- st_as_stars(restProxy3)
dim(restProxy4)

restProxy5 <- restProxy %>% filter(band==1) %>%
  st_apply(1:2, mean)
dim(restProxy5)
restProxy6 <- st_as_stars(restProxy5)
dim(restProxy6)

restProxy7 <- restProxy %>% filter(band==1) 
restProxy7 <- restProxy7 %>%  st_apply(1:2, mean)
dim(restProxy7)
restProxy8 <- st_as_stars(restProxy7)
dim(restProxy8)

```{r}
dim(temp)
st_get_dimension_values(temp, "time")
#
#tic("execute")
#
real_temp_summer2 <- st_as_stars(real_temp_summer)
real_temp_summer <- st_as_stars(real_temp_summer)

#toc()
temp <- st_as_stars(temp)

write_stars(temp, "P:/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/dummy_data/temp_test.tif")

write_stars(real_temp_summer, "P:/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/dummy_data/real_temp_summer.tif")

dim(temp)

dim(real_temp_summer)
names(real_temp_summer)

st_get_dimension_values(real_temp_summer, "time")
```

>>>>

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
#library(tidyverse)
library(dplyr)
library(ggplot2)
library(stars)
library(sf)
library(tmap)
library(lubridate)
library(ggpubr)
library(microbenchmark)
library(tictoc)
library(ncmeta)
library(parallel)
library(readr)
dir <- substr(getwd(), 1,2)

@@

myFiles <- list.files(path2, pattern=".nc$",full.names = T)
real_temp_summer <- NULL

cl <- makeCluster(10L)

@@

  tic("init")
  temp <- stars::read_ncdf(paste(myFiles[i]), var="tg", proxy=F,
                           ncsub = cbind(start = c(1, 1, julian_start), 
                              count = c(NA, NA, step)))
  year_temp <- year(st_get_dimension_values(temp, "time")[1])
  print(year_temp)
  #start <- ym(paste(year_temp, start_month, sep="-"))
  #end <-  ym(paste(year_temp, end_month, sep="-"))
  #myInterval <- interval(start, end)
  #myInterval <- interval(start, start+1)
  lookup <- setNames("mean", paste0("v_", year_temp))
  st_crs(temp) <- 25833
  toc()

  tic("filter and st_apply")
  temp <- temp %>%
    #filter(time %within% myInterval) %>%
    st_apply(1:2, mean, CLUSTER = cl) %>%
    rename(all_of(lookup)) 
  toc()

@@

tic("Merge")
real_temp_summer <- real_temp_summer %>%
  merge(name = "Year") %>%
  setNames("climate_variable")
toc()

stopCluster(cl)

@@

Here's a test for the effect of splitting over more cores.
```{r, eval=F}
cl <- makeCluster(2L)
cl2 <- makeCluster(6L)
cl3 <- makeCluster(10L)

tic("No cluster")
temp %>%
    #filter(time %within% myInterval) %>%
    st_apply(1:2, mean) %>%
    rename(all_of(lookup)) 
toc()

tic("Two clusters")
temp %>%
    #filter(time %within% myInterval) %>%
    st_apply(1:2, mean, CLUSTER = cl) %>%
    rename(all_of(lookup)) 
stopCluster(cl)
toc()

tic("Six clusters")
temp %>%
    #filter(time %within% myInterval) %>%
    st_apply(1:2, mean, CLUSTER = cl2) %>%
    rename(all_of(lookup)) 
stopCluster(cl2)
toc()

tic("Ten clusters")
temp %>%
    #filter(time %within% myInterval) %>%
    st_apply(1:2, mean, CLUSTER = cl3) %>%
    rename(all_of(lookup)) 
stopCluster(cl3)
toc()

```
No cluster: 21.401 sec elapsed
Two clusters: 21.546 sec elapsed
Six clusters: 14.548 sec elapsed
Ten clusters: 15.29 sec elapsed
Ten clusters (second run): 13.617 sec elapsed

The RStudio server has 48 cores. More parallel cores is probably on average faster. The NINA guidelines is to use max 10 cores, and to remember to close parallel cluster when done.

```{r}
#write_stars(real_temp_summer, "P:/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/dummy_data/real_temp_summer.tif")
```

I might want to merge this

```{r}
real_temp_summer <- real_temp_summer %>%
  merge(name = "Year") %>%
  setNames("climate_variable")
```

```{r}
ggplot()+
  geom_stars(data = real_temp_summer[,,,1:2], downsample = c(10, 10, 0)) +
  coord_equal() +
  theme_void() +
  scale_fill_viridis_c(option = "D") +  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0))+
  facet_wrap(~Year)
```
